### 2013-09-19 intro.tex
polish intro

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lack behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes, and only have limited capabilities to model interactions or relations.
We believe that context is one of the most important cues when it comes to classifying objects,
and therefore understanding scenes. Consequently, we target dense labeling of scenes on a fine-grained level.
The task of densely labeling image pixels into classes is called object class segmentation
or semantic segmentation.
\pagebreak

We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of \emph{context}.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will emerge soon.

We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent and process.

@@

The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We will focus on the use of conditional random fields~(CRFs), which have shown
great promise for computer vision applications. Using the paradigm of
structural support vector machines, it is possible to learn conditional random
fields to directly minimize a loss of interest. In particular, conditional
random fields allow to combine different cues, possibly produced by other
approaches, in a principled manner. One of the main difficulties with CRF
approaches to computer vision problems is, that context in images is usually
represented as a neighborhood graph.  These graphs, by nature, contain many
cycles, making inference intractable in general. Consequently, learning algorithms
have to rely on approximate inference, with often unknown consequences to learning.

@@

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where inference is simple.
More recently, \citet{lucchi2013learning} proposed a novel algorithm for
efficiently learning structured prediction for semantic segmentation, using
approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
algorithm that runs in similar time to the one proposed by
\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
the same dataset.

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introducing a general clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstrating a scalable algorithm for weakly supervised object class segmentation.
\item Providing a general and efficient software framework for structured prediction.
\item Analysing  max-margin learning algorithms with exact or approximate inference in different applications.
\item Showing that exact learning for semantic segmentation and object class
    segmentation is possible in practice, even in loopy graphs.
\end{itemize}

\section{Thesis Outline}
Before we delve into semantic segmentation and object class recognition, we
will first investigate general clustering algorithm in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. We introduce a novel, information theoretic algorithm that
compares favourably with algorithms from the literature.  While we do not apply
our algorithm to the task of bottom-up segmentation, this is a promising avenue
for future research.

We will then introducing a mostly unstructured
approach to semi-supervised learning for object class segmentation in
Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation.

@@

Chapter~\ref{ch:comparison} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

>>>>

%\includegraphics[width=\linewidth]{python_soul}%
%\flushright%
%Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lag behind
humans in most tasks that require semantic information. Often these tasks can be formulated
in terms of \emph{semantic classes}, meaning categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image?'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes, and only have limited capabilities to model interactions or relations.
We believe that \emph{context} is one of the most important cues when it comes to classifying objects,
and therefore understanding scenes. Consequently, we target dense labeling of scenes, taking
object relations into account.
The task of densely labeling image pixels into classes is called object class segmentation
or semantic segmentation.
\pagebreak

We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of context.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of manually annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will emerge soon.

In the following we distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent and process.

@@

The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.
We focus on the use of conditional random fields~(CRFs), which have shown
great promise for computer vision applications. Using the paradigm of
structural support vector machines, it is possible to learn conditional random
fields to directly minimize a loss of interest. In particular, conditional
random fields allow to combine different cues, possibly produced by other
approaches, in a principled manner. One of the main difficulties with CRF
approaches to computer vision problems is, that context in images is usually
represented as a neighborhood graph.  These graphs, by nature, contain many
cycles, making inference intractable in general. Consequently, learning algorithms
have to rely on approximate inference, with often unknown consequences to learning.

@@

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for efficient inference.
Therefore, conditional maximum likelihood learning is intractable in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where efficient inference is possible.
More recently, \citet{lucchi2013learning} proposed a novel algorithm for
efficiently learning structured prediction for semantic segmentation, using
approximate inference. In Chapter~\ref{ch:exact_learning}, we introduce an
algorithm that runs in similar time to the one proposed by
\citet{lucchi2013learning}, but is able to learn a CRF to the exact optimum on
the same dataset.

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introducing a clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstrating a scalable algorithm for weakly supervised object class segmentation.
\item Providing a general and efficient software framework for structured prediction.
\item Analysing  max-margin learning algorithms with exact or approximate inference in different applications.
\item Showing that exact learning for semantic segmentation and object class
    segmentation is possible in practice, even in loopy graphs.
\end{itemize}

\section{Thesis Outline}
Before we delve into semantic segmentation and object class recognition, we
first investigate a general clustering algorithm in Chapter~\ref{ch:itm}.
Clustering is an important step in most semantic segmentation pipelines, in at
least two places: bottom-up segmentation and creation of dictionaries for
feature learning. We introduce a novel, information theoretic algorithm that
compares favourably with algorithms from the literature.  While we do not apply
our algorithm to the task of bottom-up segmentation, this is a promising avenue
for future research.

We introducing an algorithm for  semi-supervised learning of object class
segmentation in Chapter~\ref{ch:semi_supervised}, motivated by the problem of
annotating training data for semantic segmentation.

@@

Chapter~\ref{ch:comparison} gives a quantitative comparison of the most widely
used structured prediction algorithms in diverse settings.
%, focusing on which algorithm is best to %TODO formulation
%use in a given situation.
In particular, we investigate learning behavior for
semantic segmentation on several datasets.
